Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-12 paper.qmd
changes by DC but not checked in early

```{r}
#| label: import-scripts
source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
source("scripts/five_gau_clusters/5_gen_model_with_tsne.R")
```

@@

boxplot2 <- ggplot(df_bin_centroids_two_curvy, 
       aes(x=bin_counts)) + 
  geom_histogram()

@@

hex_grid_coloured_two_curvy1 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts, 
    aes(x = x, y = y,
      group = hex_poly_id, 
      fill = std_counts), color = "grey70", linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
             aes(x = tSNE1, y = tSNE2),
             alpha = 0.3,
             size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "E") +
  interior_annotation("b", sc_ltr_pos) + 
  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1))

@@

trimesh_two_curvy <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = tsne_two_curvy_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  interior_annotation("a1", sc_ltr_pos)

@@

bin_width <- hb_obj_two_curvy$a1

####With diff benchmark values

tr_from_to_df_two_curvy4 <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3,
  by = c("from", "to"))

tr_from_to_df_two_curvy3 <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2,
  by = c("from", "to"))

## Computed benchmark value
tr_from_to_df_two_curvy <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy,
  by = c("from", "to"))

@@

```{r}
#| label: dist-plot

@@

### Compute highD distance
dist_vec <- proxy::dist(x = df_bin_two_curvy_temp[, -1], method = "Euclidean") |> as.vector()

from_vec <- c()
to_vec <- c()
num_obs <- 1:(NROW(df_bin_two_curvy_temp) - 1)

  from_val <- rep(obs, (NROW(df_bin_two_curvy_temp) - obs))
  if ((obs + 1) <= NROW(df_bin_two_curvy_temp)) {
    to_val <- (obs + 1):NROW(df_bin_two_curvy_temp)
  }
  from_vec <- append(from_vec, from_val)
  to_vec <- append(to_vec, to_val)

# Define bin width and starting point
start_point <- bin_width/2

distance_points <- ggplot(
  distance_two_curvy_dist, 
  aes(x = distance,
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 2.5 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 2 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

two_curvy_proj_umap_first_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("c", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_all_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("d", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_model_benchmark2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df2, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("b2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )
```

@@

boxplot1 <- ggplot(df_bin_centroids2, 
                   aes(x=bin_counts)) + 
    geom_histogram()

@@

@@

hex_grid_coloured_two_curvy2 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_two_curvy2, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", 
        linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
           aes(x = tSNE1, y = tSNE2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "C") +
  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1)) +
  interior_annotation("a", sc_ltr_pos)

@@

@@

@@

@@

boxplot3 <- ggplot(df_bin_centroids3, 
                   aes(x=bin_counts)) + 
    geom_histogram()

@@

error_df_two_curvy_abs <- error_df_two_curvy_abs |>
  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))

@@

error_plot_two_curvy <- error_df_two_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = sqrt_row_wise_total_error)) +
  geom_point(alpha=0.5) +
  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
  theme(
    aspect.ratio = 1
  )

error_plot_two_curvy_hist <- 
  ggplot(error_df_two_curvy_abs) +
  geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
  geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
  xlab(expression(group("|", e[hj], "|"))) +
  ylab("") +
  theme_bw() 
```

```{r}
#| label: fig-p-d-error-in-2d-two-curvy
#| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout. Color indicates square root of total error, dark blue indicating high error and light indicates low error. Most large errors are distributed near the edges of the clusters."
#| out-width: 60%
#| fig-pos: H

@@

```{r}
#| echo: false
#| label: fig-bins-two-curvy
#| fig-pos: H
#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 255 \\text{ } (15, \\text{ }17)$, (b) $b = 598 \\text{ } (23, \\text{ }26)$, and (c) $b = 2544 \\text{ } (48, \\text{ }53)$. Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins. Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
#| fig-width: 9
#| fig-height: 3

@@

hex_grid_coloured_two_curvy2 +
  hex_grid_coloured_two_curvy1 +
  hex_grid_coloured_two_curvy3 +
  #boxplot1 + boxplot2 + boxplot3 +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

@@

free(distance_points) /
  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_first_model1, ncol = 3) +
  plot_annotation(tag_levels = list(c("(a)", "(b)")))

@@

tsne_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")

nldr_two_non_linear_diff_shaped_close_clusters6 <- tsne_two_non_linear_diff_shaped_close_clusters |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("f", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

@@

error_two_non_linear_diff_shaped_close_clusters_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
error_two_non_linear_diff_shaped_close_clusters_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
error_two_non_linear_diff_shaped_close_clusters_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
error_two_non_linear_diff_shaped_close_clusters_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
error_two_non_linear_diff_shaped_close_clusters_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
error_two_non_linear_diff_shaped_close_clusters_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")

error_two_non_linear_diff_shaped_close_clusters_tsne2 <- error_two_non_linear_diff_shaped_close_clusters_tsne2 |>
  mutate(method = "tSNE2")

error_two_non_linear_diff_shaped_close_clusters <- bind_rows(error_two_non_linear_diff_shaped_close_clusters_umap, 
                         error_two_non_linear_diff_shaped_close_clusters_tsne,
                         error_two_non_linear_diff_shaped_close_clusters_phate,
                         error_two_non_linear_diff_shaped_close_clusters_trimap,
                         error_two_non_linear_diff_shaped_close_clusters_pacmap,
                         error_two_non_linear_diff_shaped_close_clusters_tsne2)

@@

error_plot_two_non_linear_diff_shaped_close_clusters <- ggplot(error_two_non_linear_diff_shaped_close_clusters, 
                           aes(x = a1, 
                               y = MSE, 
                               colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363', '#984ea3')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). Layout d is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 60%

free(error_plot_two_non_linear_diff_shaped_close_clusters) + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3, 
                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5,
                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2)
```

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 26
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

# benchmark_gau1 <- 0.1

@@

trimesh_removed_gau_tsne <- ggplot() + 
  geom_point(
    data = tsne_gau_scaled,
    aes(
      x = tSNE1,
      y = tSNE2
    ),
    alpha = 0.2) + 
  # geom_rect(
  #     aes(xmin = 0.3, 
  #         xmax = 0.65, 
  #         ymin = -0.05, 
  #         ymax = 0.35),
  #         fill = "transparent", 
  #     color = "grey50", 
  #     linewidth = 0.5) +
  interior_annotation("a1", 
                      position = c(0.08, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

axes_obj <- gen_axes(  
  proj = projection,
  limits = 1,
  axis_pos_x = -0.6,
  axis_pos_y = -0.63,
  axis_labels = names(scaled_gau_data),
  threshold = 0)

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.3,
    color = "#999999") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.84, 0.84)) +
  ylim(c(-0.84, 0.84)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

>>>>

```{r}
#| label: import-scripts
source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
```

@@

@@

hex_grid_coloured_two_curvy1 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts, 
    aes(x = x, y = y,
      group = hex_poly_id, 
      fill = std_counts), color = "grey70", linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
             aes(x = tSNE1, y = tSNE2),
             alpha = 0.3,
             size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "E") +
  interior_annotation("a", sc_ltr_pos) + 
  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1))

@@

@@

tr_from_to_df_two_curvy <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy,
  by = c("from", "to"))

@@

@@

@@

two_curvy_proj_umap_first_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("c", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_all_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("d", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

@@

@@

### Compute highD distance
dist_vec <- proxy::dist(x = df_bin_two_curvy2[, -1], method = "Euclidean") |> as.vector()

from_vec <- c()
to_vec <- c()
num_obs <- 1:(NROW(df_bin_two_curvy2) - 1)

  from_val <- rep(obs, (NROW(df_bin_two_curvy2) - obs))
  if ((obs + 1) <= NROW(df_bin_two_curvy2)) {
    to_val <- (obs + 1):NROW(df_bin_two_curvy2)
  }
  from_vec <- append(from_vec, from_val)
  to_vec <- append(to_vec, to_val)

@@

hex_grid_coloured_two_curvy2 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_two_curvy2, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", 
        linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
           aes(x = tSNE1, y = tSNE2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "C") +
  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1)) +
  interior_annotation("b", sc_ltr_pos)

@@

# Define bin width and starting point
bin_width <- hb_obj_two_curvy2$a1
start_point <- bin_width/2

distance_points <- ggplot(
  distance_two_curvy_dist, 
  aes(x = distance,
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  #geom_jitter() +
  # scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
  #                    labels = c(
  #                      expression(a[1]),
  #                      expression(2*a[1]),
  #                      expression(3*a[1]),
  #                      expression(4*a[1]),
  #                      expression(5*a[1]),
  #                      expression(6*a[1]),
  #                      expression(7*a[1]),
  #                      expression(8*a[1])
  #                    )) +
  geom_vline(xintercept = 2.5 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 2 * bin_width,
           linetype="solid",
           color = "#bdbdbd",
           linewidth=1) +
  # geom_vline(xintercept = 4.5,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  #ggtitle("(a)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

distance_two_curvy_dist_summary <- distance_two_curvy_dist |>
  count(bin) |>
  mutate(prop = n/sum(n)) |>
  head(4)

# distance_two_curvy_dist <- distance_two_curvy_dist |>
#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small

text_df <- tibble(
  x = c(1.9, 3.4),
  y = c(0.5, 0.5),
  text = c("e1", "d1")
)

distance_hist <- ggplot(
  distance_two_curvy_dist_summary,
  aes(x = as.numeric(bin),
      y = prop)) +
  geom_line() +
  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin)))[1:4],
                     labels = c(
                       expression(a[1]),
                       expression(2*a[1]),
                       expression(3*a[1]),
                       expression(4*a[1])
                     )) +
  geom_vline(xintercept = 2,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 3.5,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_text(
    data = text_df,
    aes(
      x = x,
      y = y,
      label = text
    ),
    size = 9
  ) +
  ylab("Proportion of the number of edges") +
  xlab(expression(d^{(2)})) +
  ggtitle("(b)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

# distance_hist <- ggplot(
#   distance_two_curvy_dist, aes(x = distance)) +
#     geom_histogram(
#       aes(y = after_stat(count) / sum(after_stat(count))), 
#       binwidth = bin_width) +
#   scale_x_continuous(breaks = bin_labels, 
#                      labels = c(
#                        expression(a[1]),
#                        expression(2*a[1]),
#                        expression(3*a[1]),
#                        expression(4*a[1]),
#                        expression(5*a[1]),
#                        expression(6*a[1]),
#                        expression(7*a[1]),
#                        expression(8*a[1])
#                      )) +
#   geom_vline(xintercept = start_point + 2*bin_width,
#              linetype="solid",
#              color = "#bdbdbd",
#              linewidth=1) +
#   ylab("Proportion of the number of edges") +
#   xlab(expression(d^{(2)})) +
#   #ggtitle("(d)") +
#   theme_minimal() +
#   theme(aspect.ratio = 0.75,
#         panel.border = element_rect(fill = 'transparent'),
#         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
#         panel.grid.major.x = element_blank(),
#         axis.ticks.x = element_line(),
#         axis.ticks.y = element_line())

@@

tr_from_to_df_two_curvy3 <- inner_join(
  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
  by = c("from", "to"))

tr_from_to_df_two_curvy4 <- inner_join(
  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy3,
  by = c("from", "to"))

@@

two_curvy_proj_umap_model_benchmark2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df2, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("b2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

@@

@@

error_df_two_curvy_abs <- error_df_two_curvy_abs |>
  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))

@@

error_plot_two_curvy <- error_df_two_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = sqrt_row_wise_abs_error)) +
  geom_point(alpha=0.5) +
  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
  theme(
    aspect.ratio = 1
  )

error_plot_two_curvy_hist <- 
  ggplot(error_df_two_curvy_abs) +
  geom_histogram(aes(x=sqrt_row_wise_abs_error^2, y=..density..)) +
  geom_density(aes(x=sqrt_row_wise_abs_error^2, y=..density..), colour="red") +
  xlab(expression(group("|", e[hj], "|"))) +
  ylab("") +
  theme_bw() 
```

```{r}
#| label: fig-p-d-error-in-2d-two-curvy
#| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout. Color indicates square root of absolute error, dark blue indicating high error and light indicates low error. Most large errors are distributed near the edges of the clusters."
#| out-width: 60%
#| fig-pos: H

@@

```{r}
#| echo: false
#| label: fig-bins-two-curvy
#| fig-pos: H
#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$. Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins. Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
#| fig-width: 9
#| fig-height: 3

@@

hex_grid_coloured_two_curvy2 +
  hex_grid_coloured_two_curvy1 +
  hex_grid_coloured_two_curvy3 +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

@@

free(distance_points) /
  wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_model, ncol = 3) +
  plot_annotation(tag_levels = list(c("(a)", "(b)")))

@@

@@

error_two_non_linear_diff_shaped_close_clusters_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
error_two_non_linear_diff_shaped_close_clusters_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
error_two_non_linear_diff_shaped_close_clusters_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
error_two_non_linear_diff_shaped_close_clusters_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
error_two_non_linear_diff_shaped_close_clusters_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")

error_two_non_linear_diff_shaped_close_clusters <- bind_rows(error_two_non_linear_diff_shaped_close_clusters_umap, 
                         error_two_non_linear_diff_shaped_close_clusters_tsne,
                         error_two_non_linear_diff_shaped_close_clusters_phate,
                         error_two_non_linear_diff_shaped_close_clusters_trimap,
                         error_two_non_linear_diff_shaped_close_clusters_pacmap)

@@

error_plot_two_non_linear_diff_shaped_close_clusters <- ggplot(error_two_non_linear_diff_shaped_close_clusters, 
                           aes(x = a1, 
                               y = MSE, 
                               colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

```{r}
#| fig-cap: "Assessing which of the 5 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). Layout d is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 60%

free(error_plot_two_non_linear_diff_shaped_close_clusters) + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3, 
                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5, ncol = 2)
```

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 20 #13
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

benchmark_gau1 <- 0.1

@@

trimesh_removed_gau_tsne <- ggplot() + 
  #geom_segment(data = tr_from_to_df_gau1, 
  #             aes(
  #               x = x_from, 
  #               y = y_from, 
  #               xend = x_to, 
  #               yend = y_to),
  #             colour = "#000000",
  #             linewidth = 1) +
  geom_point(
    data = tsne_gau_scaled,
    aes(
      x = tSNE1,
      y = tSNE2
    ),
    alpha = 0.2) + 
  geom_rect(
      aes(xmin = 0.3, 
          xmax = 0.65, 
          ymin = -0.05, 
          ymax = 0.35),
          fill = "transparent", 
      color = "grey50", 
      linewidth = 0.5) +
  interior_annotation("a1", 
                      position = c(0.08, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

axes_obj <- gen_axes(  
  proj = projection,
  limits = 0.3,
  axis_pos_x = 0.05,
  axis_pos_y = -0.85,
  axis_labels = names(scaled_gau_data),
  threshold = 0)

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 1) +
  geom_point(
    #size = 0.5,
    alpha = 0.2,
    color = "#999999") +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(0, 0.4)) +
  ylim(c(-0.9, -0.5)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)
